### 2019-03-22 eDriversMS-SI.rmd

---
title: Supplementary Material
fontsize: 12pt
output:
  pdf_document:
    toc: false
    number_sections: true
  html_document:
    number_sections: true
relativeurls: true
bibliography: eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

<!--
cd phdproj/edrivers/edriversms
rmarkdown::render('./eDriversMS-SI.rmd', 'pdf_document')
rmarkdown::render('./eDriversMS-SI.rmd', 'html_document')
-->

# Drivers description

Table S1. List of drivers currently available on *eDrivers* along with their respective accronym used in the figures in the supplementary material.

```{r tableS1, echo=F, fontsize=4}
load('./tables/tableS1.RData')
tableS1
```

## Climate

### Aragonite

***Describe data***

We interpolated aragonite distribution using the exponential kriging model.
We then built an index of aragonite level stress. When $\Omega$ aragonite levels
decrease below 1, the water becomes corrosive to calcifying organisms such as
echinoderms, corals and mollusks larvae. These organisms thus have to
continually recalcify their shells to maintain their structure, which becomes
harder to do as $\Omega$ aragonite levels further decrease towards 0. We therefore
considered stress associated to decreasing $\Omega$ aragonite levels to increase
linearly between values of 1 and 0. The index of stress therefore has a value of
0 and 1 for values of $\Omega$ aragonite equal to 1 and 0, respectively.

### Hypoxia

***Describe data***

We interpolated the dissolved oxygen using cokriging with depth as a covariable,
as done in [@dutil2011]. According to @diaz1995, severe hypoxia can be observed
when oxygen saturation falls below 2 $ml$ $L^{-1}$ or 62.5 $\mu mol$ $L^{-1}$, which is considered
the level necessary to maintain most animal life. This threshold is equal to 1.4
$ml$ $L^{-1}$ (1 $mg$ $L^{-1}$ = 0.7 $ml$ $L^{-1}$).
We used this threshold to create an index of hypoxia using an inverted logistic curve
to transform the dissolved oxygen values as an index of hypoxic stress $H_s$ ranging
from 0 to 1, under the assumption that hypoxic stress will increase following
a logistic curve as it approaches the hypoxic threshold:

$$H_s = \frac{-1}{200 * e^{-2.5 * DO_2}} + 1$$

### Bottom temperature anomalies

The data used to characterize bottom temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods
to characterize surface temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

Bottom temperatures are interpolated in the Gulf using
conductivity-temperature-depth (CTD) sampling performed annually through DFO's
multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. Using this sampling survey, temperatures are interpolated
at each 1 m depth layer on a 2 km resolution grid. Bottom temperature are then
extracted from this grid by using a bathymetry layer from the Canadian
Hydrographic Survey [@dutil2012].

We used temperature anomalies, *i.e.* deviations from long-term normal conditions,
to measure an annual index of stress associated with extreme temperatures between
2013 and 2017. Temperature anomalies were calculated using the difference between
grid cell monthly climatologies to the associated long-term averages generated
from the reference period between 1981 and 2010. Grid cells whose monthly value
exceeded ±0.5 standard deviation (SD) from the long-term average were considered
as anomalous [@galbraith2018].

Outliers in the data were considered as those that fell beyond the
interquartile range (IQR) * 3, identified as extreme ouliers by @tukey1977.
Outlier values were capped to correspond to the 5th and 95th percentiles.
Anomalies were divided into positive and negative anomalies and
summed for the available months of September and August.
By adding the monthly anomaly values, we essentially considered the summed
deviation from the mean as an indicator of the annual intensity of surface
temperature anomalies in the St. Lawrence.

### Surface temperature anomalies

The data used to characterize surface temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods
to characterize surface temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

The surface layer is characterized using a variety of methods:

1. Temperature-salinity sensors installed on various sihps and forming the
shipboard thermosalinograph network,
2. Moored instruments recording water temperature every 30 minutes and forming
the thermograph network,
3. Sea surface temperature (SST) monthly composite climatologies using Advanced
Very High Resolution Radiometer (AVHRR) satellite images from the National
Oceanic and Atmospheric Administration (NOAA) and European Organisation for the
Exploitation of Meteorological Satellites (EUMETSAT) available from the Maurice
Lamontagne Institute sea surface temperature processing facility at a 1km
resolution from 1985-2013 and from the Bedford Institute of Oceanography (BIO)
Operational Remote Sensing group at a 1.5km resolution since 2014.

Positive and negative surface temperature anomalies were identified following
the same approach used for bottom temperature anomalies. Compared to the
bottom temperature anomalies, however, data is available thoughout the year.
Only the months of April to November were included to avoid biases associated
with anomalies measurements due to the presence of ice cover.

### Sea level rise

The data used to characterize sea level rise risk comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Sea level rise was characterized by @nicholls2010 using NASA's satellite
altimetry data (Topex/Poseidon, Jason-1&2, GFO, ERS-1&2, and Envisat
missions) and available at
http://www.aviso.altimetry.fr/en/data/products/ocean-indicatorsproducts/mean-sea-level/products-images.html

The rate of sea level rise ($mm/year$) was measured between 1992 and 2012 and
transformed as a net change value ($mm$) by multiplying by the number of years
considered. Only positive values were selected under the assumption that
only positive sea level rise is likely to cause environmental stress.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

## Coastal

### Aquaculture

Aquaculture data comes from a variety of sources in the St. Lawrence because
aquaculture sites are mostly managed at the provincial level. We therefore had
to gather the data on aquaculture sites from the 5 provinces dividing the
St. Lawrence.

Invertebrates aquaculture is especially important in the southern and western
Gulf. Fish and algae aquaculture, on the other hand, remains marginal.
Considering this, we only considered invertebrates aquaculture for the
aquaculture driver layer. However, if fish or algae farming were to become more
important, these driver should be incorporated in future analyses as individual
layers, as impacts vary between types of aquaculture.

Aquaculture activities are highly localized and potential effects do not or
rarely extend beyond the location of the farms. We therefore only considered
the actual location of sites to characterize the distribution of this driver.
We were unable to characterize site production in terms of biomass farmed, which
could provide an indication of the intensity of aquaculture activities. As such,
we considered aquaculture as binary presence-absence data in our analyses.

### Coastal development

We used lights at night as a proxy of coastal infrastructure development, as
terrestrial stable lights at night represent light from human settlements and
industrial sites with electricity.

The data comes from the Nighttime Lights Time Series. Nighttime light products
are compiled by the Earth Observation Group at the National Oceanic and
Atmospheric Administration's (NOAA) National Centers for Environmental
Information (NCEI). They use glogally available nighttime data obtained from
the Visible Infrared Imaging Radiometer Suite (VIIRS) Day/Night Band (DNB) of
the Defense Meteorological Satellite Program (DMSP) to characterize global
average radiance ($nanoWatts$ $cm^{-2}$ $sr^{-1}$) composite images at a
15 arc-second (~200 m) resolution.

We used the annual Version 1 Nighttime VIIRS DNB composites between 2015 and
2016 [@eog2019] to characterize coastal development in coastal areas of the
St. Lawrence. As the effects of coastal development are likely acute in its
direct vicinity, we extracted average radiance values using a 2 km buffer
around grid cells within 2 km of the coast. We used a weighted area average to
extract the radiance values.

### Direct human impact

As in @halpern2008 and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statscan2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated and they combine
to cover all of Canada. The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

As the effects of direct human impacts are likely acute mostly in coastal areas
we calculated total population in grid cells within 2 km of the coast. Total
population was measured in a 10 km buffer around each coastal cell.
The total population in each buffer was the sum of intersecting dissemination
areas divided by the intersection area between buffers and dissemination areas:

$$DHI_j = \sum_{k=1}^{n_j} P_k * \frac{A_{j,k}}{A{tot, k}}$$

where $j$ is a buffered grid cell, $k$ is a dissemination area intersecting $j$,
$P$ is the population in $k$, $A$ is the area of the $k$ overlapping with $j$
and $A_{tot}$ is the total area of $k$. This approach was favoured to reduce the
effects of very large dissemination areas overlapping with buffers on a very
small percentage of their total area.

### Inorganic pollution

The data used to characterize inorganic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Inorganic pollution was modelled using impervious surface area (*i.e.*
artificial surfaces such as paved roads) under the assumption that most of this
pollution source comes from urban runoff. Inorganic pollution originating from
point-sources or in areas lacking paved roads is therefore not captured by
this layer. The data obtained was aggregated at the watershed scale and spread
into coastal and marine environments was modelled using a diffusive plume model
from each watershed pourpoints (*e.g.* river mouths).

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

### Nutrient pollution

The data used to characterize nutrient pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual fertilizer use in tonnes ($t$) was used as a proxy of nutrient pollution.
The data used came from the Food and Agriculture Organization of the United
Nations (FAO). Gaps in data were modelled using a linear regression between
fertilizer and pesticides or agricultural gross domestic product (GDP).
Dasymetric maps were then used to distribute fertilizer data over the landscape
using 2009 data from the Moderate Resolution Imaging Spectroradiometer (MODIS)
at ~500 m resolution and aggregated to watersheds. Diffusive plume models
from each watershed pourpoint (*e.g.* river mouths) were then used to model the
distribution and intensity of nutrient pollution in coastal and marine
environments.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

### Organic pollution

The data used to characterize organic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual pesticide use in tonnes ($t$) was used as a proxy of organic pollution.
The data used came from the FAO and gaps in data were modelled using a linear
regression between pesticides and fertilizers or agricultural GDP. The same
methodology as that used to characterized nutrient pollution was then applied
to organic polllution.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

### Toxic algae

The data we use to describe the risk of toxic algae comes from an expert based
map delineating the areas where coastal areas are at risk from five different
toxins [@bates2019]. The map presents coastal areas at risk from 5
different toxins: 1) paralytic shellfish poisoning (PSP) toxins from the regular
presence of the dinoflagellate *Alexandrium catenella* (previously
*Alexandrium tamarense*) at high concentrations, 2) amnesic shellfish
poisoning (ASP) toxins from domoic acid 3) diarrhetic shellfish poisoning (DSP)
toxins, 4) spirolides and 5) pectenotoxins,  two toxins produced by
dinoflagellates occurring in the St. Lawrence.

The information provided on this expert map on the 5 toxins [@bates2019].
was georeferenced and transformed as vecctorized objects.
We calculated a toxic algae risk ($T$) index for each cell ($x$) in the 1 $km^2$
study grid. For each toxin ($t$), a value of 1 was attributed to all grid cells
overlapping with areas identified at risk on the expert map and a value of 0.5
for grid cells overlapping with areas where ASP and DSP toxins were observed
without exceeding legal thresholds. The value for all 5 toxins was them summed
for all grid cells:

$$TA_{i,x} = \sum_{i = 1}^{5} i_x$$

## Fisheries

The impacts of fisheries activities in the St. Lawrence are evaluated using
DFO's fisheries logbooks program [@dfo2016]. While logbooks are not mandatory
for all fisheries in the St. Lawrence, they still provide a very thorough
overview of the spatial distribution and intensity of fishing activities in the
St. Lawrence. The data we used spans 6 years from 2010 to 2015 and details
218323 fishing events (36387 $\pm$ 3147 fishing events per year). There were 31
targetted species and a total of 53 caught species in the dataset.

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008 and
@halpern2015b and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization therefore divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

Gear types can also be further classified into fixed or mobile engines based
on their mobility. We used these two mobility classes to generate a buffer of
impact around each fishing activity coordinates to consider potential spatial
uncertainty associated with locations and the fact that mobile engines can be
tracted over several kilometers during fishing activities and that we do not
have the beginning and end points of mobile fishing events. Buffer sizes for
fixed and mobile engine was of 200 and 2000 meters, respectively.

Table S2. Classification of gear types in the fisheries dataset based on their
environmental impact and mobility

Gear type (EN)    | Classification    | Mobility
--------------    | --------------    | --------
Trap              | DNH               | Fixed
Trawl             | DD                | Mobile
Dredge            | DD                | Mobile
Driftnet          | PHB               | Fixed
Hand lines        | PLB               | Fixed
Longline          | PHB               | Fixed
Scuba diving      | DNL               | Fixed
Purse seine       | PLB               | Fixed
Seine             | DNH               | Fixed
Beach seine       | DNH               | Fixed
Trap              | DNH               | Fixed
Jig fishing       | PLB               | Fixed

In order to characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multipled the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

$$FI_j = \sum_{k=1}^{n_j} B_{tot, k} * \frac{A_{j,k}}{A_{tot,k}}$$

where $j$ is a grid cell, $k$ is a fishing event, $B_{tot}$ is the total biomass
of a fishing event $k$, $A$ is the area of a fishing event *k* overlapping a
cell $j$ and $A_{tot}$ is the total area of the fishing event $k$. This formula
gives an intensity measurement in biomass units, which is kg in our case. Since
we measure the intensity within a 1 $km^2$ grid cell, the intensity evaluation
is in $kg * km^{-2}$. This metric distributes the biomass captured within each
grid cell as a function of overlapping fishing area and provides an overview of
how impacted each grid cell is in terms of extracted biomass.

## Marine traffic

### Shipping

The data used to characterize shipping comes from the global cumulative impacts
assessment on habitats [@halpern2008; @halpern2015] and available on the NCEAS
online data repository [@halpern2015b]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

Two data sources were used to characterize shipping. The first set of data is
gathered as part of the World Meteorological Organization Voluntary Observing
Ships' (VOS) scheme [@jcomm2018]. Ships participating in the program gather
meteorological data along with observation location as part of an open-ocean
climate dataset. The data spans 20 years and annually covers 10-20% of ships
worlwide. Data used spanned 2003 to 2011.

The second set of data comes from the Automatic Identification System (AIS),
an initiative launched in 2002 that sought to improve marine safety by providing
mariners with real-time vessel traffic [@tetreault2002]. Through the
International Maritime Organization SOLAS agreement, all vessels of over 300
gross tonnage on international voyages and those carrying passengers are now
required to be equipped with AIS transceivers. These transceivers use Global
Positioning System technology to locate vessels every 10 minutes. The data used
was from November 2010 to December 2011.

Data used come mostly from vessels that move globally (*i.e.* cargo, tanker and
passenger), as they are required to carry AIS transceivers, but also include
data from fishing, high-speed, pleasure and support classes. Shipping intensity
was evaluated as the number of fishing tracks at a 0.1 decimal degrees resolution.
For more details on data and methods used, consult [@walbridge2013].

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

### Invasive species

The data used to characterize invasive species risk comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
supplementary materials provided in @halpern2008 and @halpern2015.

Cargo volume was used as a proxy of invasion risk under the assumption that
risk of invasion is proportional to tonnes of goods transferred through ports.
Cargo throughput in metric tonnes for the year 2011 was accessed through a
variety of sources [see supplementary material in @halpern2015 for more details]
and cross-matched with entries in the World Port Index database (WPI; available
from the National Geospatial-Intelligence Agency). A gap-filling procedure
using linear regression and sets of predictors related to port volume and
available in the WPI dataset was then applied to the WPI dataset to predict
missing cargo volume entries. Finally, volume data was distributed in marine
environments adjacent to ports using a diffusive plume model with an exponential
decay function that set the maximum spread distance to approximately 1000 km.
The plume model was then clipped to areas less than 60 m deep, as
invasives species are more likely to invade shallow areas.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

### Marine pollution

The data used to characterize marine pollution risk comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b].
Marine pollution was considered to be mainly driver by the shipping industry.
As such, the driver layer was constructed by combining the shipping (*i.e.*
shipping lanes) and invasives species (*i.e.* cargo volume) layers.
invasive. For more details, refer to supplementary materials provided in
@halpern2008 and @halpern2015.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

# Driver intensity and distribution

We evaluated the frequency distribution of each drivers to verify whether data
should be transformed (Figure S\ref{hist}). In light of this, we log-transformed
the following driver layers

- Coastal development
- Direct human impact
- All fisheries data
- Hypoxia
- Inorganic pollution
- Invasive species
- Nutrient pollution
- Organic pollution
- Sea bottom temperature anomalies
- Shipping

To allow for relative intensity comparison, all driver layers were subsequently
normalized between 0 and 1 using the 99th quantile to further control for
extreme values (Figure S\ref{drivers}).

![Frequency distribution of the untransformed data for all driver layers. \label{hist}](./figures/drHist.png)

![Distribution and intensity of transformed and normalized drivers in the Estuary and Gulf of St. Lawrence available on *eDrivers*. \label{drivers}](./figures/drivers.png)

\newpage

# Cumulative exposure

![Frequency distribution of cumulative exposure (*i.e.* sum of normalized driver intensity in each grid cell) and percent contribution of each driver to the frequency distribution of cumulative exposure in the Estuary and Gulf of St. Lawrence \label{marimekko}](./figures/marimekko.png)

# Threat complexes

## Clustering

We identified regions with similar cumulative exposure regimes, referred to as
threat complexes after [@bowler2019], using a partional *k-medoids* clustering
algorithm, CLARA [CLustering for Large Applications; @kaufman1990], which was
designed to use with large datasets. The CLARA algorithm uses the PAM
(Partition Around Medoids) algorithm on a sample from the original dataset
to identify a set of $k$ objects that are representative of all other objects,
*i.e.* medoids (also referred to as examplars) and that are central to the
cluster they represent. The goal of the algorithm is to iteratively minimize
dissimilarity between objects in the cluster and the representative medoid.
Sets of $k$ medoids are thus selected and replaced in order to iteratively
minimize the intra-cluster dissimilarity. With the CLARA algorithm, this
process is repeated and iterations are compared on the basis of their average
dissimilarity between objects and cluster medoids to select the optimal set of
$k$ medoids that minimizes average dissimilarity. We performed the analyses for
a total of 100 iterations using samples of 10000 observations, representing ~5%
of the drivers dataset. Analyses were performed using the *cluster* R package
[@maechler2018].

Because the algorithm identifies clusters by minimizing dissimilarity around a
set of *k* medoids, it requires a pre-defined number of clusters. The
appropriate number of clusters $k$ was thus tested using values ranging from
2 to 10 and validated by selecting the number of clusters that maximized the average
silhouette width [@kaufman1990] and minimized the total within-cluster sum of
squares (WSS; Figure S\ref{valid}).

We also validated the clustering by comparing *k-medoids* clustering with
*k-means* clustering with the *Lloyd* algorithm [@lloyd1982]. The *k-means*
approach is similar to the *k-medoids*, but identifies observations belonging
to a cluster iteratively by minimizing the mean intra-cluster squared distance
until it converges to an optimal solution. We used 25 random sets and set a
maximum of 1000 iterations for the analysis. Analyses were performed using
the *stats* R package [@rcoreteam2018]. We used the same validation procedure
to select the optimal number of clusters $k$ than with the *k-medoids*
clustering (Figure S\ref{valid}).

Both the *k-medoids* and *k-means* clustering validation procedure suggest that
there are 7 distinc threat complexes in the St. Lawrence (Figure S\ref{valid}).
While the comparison of of the spatial distribution of clusters results in only
a 66% cell correspondance between the *k-medoids* and *k-means* approaches, the
bulk of the difference is attributable to a single undivided cluster that
combines most of the southern and norther Gulf in a single cluster in favor of
a more localised cluster by the *k-means* algorithm (Figure S\ref{comparison}).
While *k-means* algorithms are much faster than *k-medoids* since it does not
need a pairwise dissimilarity measurement of all observations, it is also more
sensitive to outliers since it focuses on the mean rather than a centroid. We
therefore opted the use of *k-medoids* clustering.

![Validation procedure for the *k-medoids* and *k-means* clustering algorithms based on the number of cluster that maximizes average silhouette width [upper panels; @kaufman1990] and minimizes the total within-cluster sum of squares (WSS; lower panels). \label{valid}](./figures/validation.png)

![Comparison of the distribution of the 7 clusters identified using the *k-medoids* and *k-means* clustering algoriths and the percent grid cell correspondance between the two approaches \label{comparison}](./figures/clusterComp.png)

## Inter-cluster dissimilarity

The difference between clusters was explored by measuring the total inter-cluster
dissimilarity and the contribution of each driver to the total inter-cluster
dissimilarity using a similarity percentage analysis (SIMPER) with Bray-Curtis
dissimilarity [Figure S\ref{inter}; @clarke1993]. As the drivers dataset is too
large, we used a bootstrap procedure for the SIMPER analysis, randomly selecting
5% of each cluster to run the analysis and repeating the process over 300 iterations.
We also compared the mean intensity of each driver within each cluster to
better capture the inter-cluster dissimilarity. Analyses were performed using
the *vegan* R package [@oksanen2018].

![Evaluation of inter-cluster dissimilarity using a similarity percentage analysis (SIMPER) with Bray-Curtis dissimilarity [@clarke1993]. The figure diagonal presents the distribution of the 7 clusters identified using the *k-medoids* clustering algorithm. The lower triangle shows all combinations of inter-cluster dissimilarity with circular barplots showing the percent contribution to total dissimilarity of each driver and with the total inter-cluster dissimilarity in the center of the barplots. The upper triangle shows the average relative intensity of each driver for all driver combinations, with barplots to the left and the right representing the row and columns clusters, respectively. \label{inter}](./figures/interDissimilarity.png)

## Intra-cluster similarity

Intra-cluster similarity was evaluated using the Bray-Curtis similarity index
(Figure S\ref{intra}). As with the inter-cluster dissimilarity, we used a
bootstrap procedure for the intra-cluster similarity, randomly selecting 5% of
each cluster observation to run the analysis and repeating the process over 300
iterations. We however did not use the bootstraping procedure for clusters 2
and 6, as they both contain less than 10000 observations, making the calculation
time more manageable.

![Evaluation of intra-cluster similarity using the Bray-Curtis similarity index. The distribution of all 7 clusters is presented along with circular barplots showing the percent contribution to total similarity of each driver and with the total intra-cluster similarity in the center of the barplots. \label{intra}](./figures/intraSimilarity.png)

# References