### 2019-03-22 eDriversMS.Rmd

---
title: Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^5^, Simon Bélanger^4^, Pascal Bernatchez^4^, Hugo Bourdages^7^, Clément Chion^6^, Peter S. Galbraith^7^, Benjamin Halpern^8,9,10^, Camille Lavoie^2^, Christopher W. McKindsey^7^, Alfonso Mucci^11^, Simon Pineault^12^, Michel Starr^7^, Anne-Sophie Ste-Marie^5^, Philippe Archambault^2^
date: 2019-02-06
fontsize: 12pt
output:
  pdf_document:
    toc: false
    number_sections: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   # - \usepackage{titling}
   # - \pretitle{\begin{flushleft}}
   # - \posttitle{\end{flushleft}}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
relativeurls: true
bibliography: eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

<!--
cd phdproj/eDrivers/eDriversMS/
rmarkdown::render('./eDriversMS.rmd', 'pdf_document')
rmarkdown::render('./eDriversMS.rmd', 'word_document')
-->

<!--
<style type="text/css">
/* body{
  font-size: 12px;
}
td {
  font-size: 8px;
} */
h1 {
  font-size: 38px;
  color: DarkRed;
}

Style specs
Title{
  font-size: 16
  font: times new roman
  align: center
}
Headings & subheadings{
  font-size: 12
  font: times new roman
  bold
  numbered
}
Abstract{
  font-size: 12
  font: times new roman
  normal
}
Body{
  font-size: 12
  font: times new roman
  single-spaced
  page numbers
  line numbers
}

</style>
-->

<!-- Setting up parameters for code snippet presented in Box 1 -->
\lstset{language=R,
    basicstyle=\small\ttfamily,
    stringstyle=\color{DarkGreen},
    otherkeywords={fetchDrivers, install_github, getData, kde2d},
    morekeywords={TRUE,FALSE},
    deletekeywords={data,frame,length,as,character},
  keywordstyle=\color{blue},
  commentstyle=\color{DarkGreen},
}

^1^Institut des sciences de la mer, Université du Québec à Rimouski, Rimouski, QC, Canada \newline
^2^Québec Océan, Département de biologie, Université Laval, Québec, QC, Canada \newline
^3^Département de biologie, Université de Sherbrooke, Sherbrooke, QC, Canada \newline
^4^Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \newline
^5^St. Lawrence Global Observatory, Rimouski, QC, Canada \newline
^6^Institut des sciences de la forêt tempérée, Université du Québec en Outaouais, Gatineau, QC, Canada \newline
^7^Fisheries and Oceans Canada, Maurice Lamontagne Institute, Mont-Joli, QC, Canada \newline
^8^National Center for Ecological Analysis and Synthesis, University of California, Santa Barbara, CA, United States \newline
^9^Bren School of Environmental Science and Management, University of California, Santa Barbara, CA, United States \newline
^10^Imperial College London, Ascot, United Kingdom \newline
^11^Department of Earth & Planetary Sciences, McGill University, Montréal, QC, Canada \newline
^12^Ministère Environnement et Lutte contre les changements climatiques, Québec, QC, Canada  \newline

**Correspondence**: \newline
David Beauchesne \newline
david.beauchesne@uqar.com \newline

**Number of figures**: 5 \newline
**Number of tables**: 1 \newline

\newpage

# Abstract

Gathering environmental data for large scale, systematic initiatives such as
ecosystem-based management is a very challenging process. Building such
holistic datasets is made increasingly possible through initiatives like the
Ocean Biogeographic Information System (OBIS) and Bio-ORACLE. Data on
environmental drivers (*e.g.* climate change and fisheries) are however largely
missing even though their characterization is crucial to a better understanding
of the effects of multiple environmental threats. In this paper we discuss how
we are addressing this issue by launching an open data platform called *eDrivers*, focused
on characterizing and sharing data on the distribution and intensity of drivers
and using the Estuary and Gulf of St. Lawrence in eastern Canada as a case study.
The core objective underlying this initiative is to create a community-led
project built on a series of guiding principles aiming at upholding the highest
standards of data management and open science. The platform currently shares data for 22 drivers
including coastal, climate, fisheries and marine traffic related layers. Using
the currently available data, we show that few areas are free of cumulative
exposure in the St. Lawrence. The Estuary, the Anticosti Gyre and coastal areas
are particularly exposed, especially in the vicinity of urban centers. We also
identify 7 distinct areas of similar cumulative exposure regime (*i.e.* threat
complexes) that show that numerous drivers typically co-occur in different
regions of the St. Lawrence and that coastal areas are exposed to all driver
types. These observations are destined to improve as *eDrivers* evolves
through time to address gaps in knowledge, incorporates additional driver
layers and refines the characterization of current driver layers. Ultimately,
we believe that *eDrivers* represents a much needed solution that could
radically influence broad scale research and management practices by increasing
data accessibility and interoperability and by increasing research and
decision-making efficiency. All that is required is a commitment to open data,
adaptive monitoring and an integrated vision of ecosystem management.

**Keywords**: ocean observing systems, St. Lawrence, environmental drivers,
cumulative exposure, threat complex, multiple stressors, global change

# Introduction

With awareness of human impacts on the oceans increasing globally, there has
been a growing demand for systematic and large scale environmental management
schemes, such as ecosystem-based management, systematic conservation planning,
and strategic environmental assessment [@margules2000; @jones2016]. Beyond
substantial political and scientific commitment, such initiatives require the
proper infrastructures and tools allowing for efficient data reporting and
adaptive, continuous monitoring, all of which hinge on the construction of
large, holistic datasets.

Gathering environmental data for large scale, systematic initiatives can however
be a very challenging and time consuming - not to say painful - process. On one
hand, there is an overwhelming and expanding wealth of data available. Such
information overload may reduce ability to take decision based on scientific
information, cause massive effort duplication, disproportionately appropriate
research funds for certain sectors, and obscure knowledge gaps amid a sea of information
[@eppler2004]. On the other hand, crucial data are lacking and remain largely
unavailable or inaccessible for a variety of reasons, including proprietary
rights, lack of organizational time, capacity and training, and unwillingness
to share, curtailing our ability for appropriate decision-making. When coupled
with political intricacies, such as multi-jurisdictional management, the data
gathering process can quickly become intractable.

Moreover, the current digital infrastructure is highly
decentralized and the practices are heterogeneous, preventing us from maximizing
benefits from research investments [@wilkinson2016]. Per @wilkinson2016, *"the
reason that we often need several weeks (or months) of specialist technical
effort to gather the data necessary to answer such research questions is not
the lack of appropriate technology; the reason is, that we do not pay our
valuable digital objects the careful attention they deserve when we create and
preserve them"*.

There now exists multiple initiatives that address this issue by assembling,
organizing and sharing vast arrays of environmental data.
Biotic data can be accessed through web portals such as the Ocean Biogeographic
Information System [OBIS; @obis2018], the Global Biodiversity Information Facility
[GBIF; @gbif2018], the Global Biotic Interactions platform [GloBI; @poelen2014],
and the World Register of Marine Species [WoRMS; @worms2017]. Abiotic data can
also be accessed through WorldClim [@hijmans2005], Bio-ORACLE
[@tyberghein2012], and MARSPEC [@sbrocco2013]. Initiatives focused on sharing
environmental data for specific areas also exist, such as the St. Lawrence
Global Observatory (SLGO; https://ogsl.ca/en), the European Marine Observation
and Data Network (EMODnet; http://www.emodnet.eu/) and the U.S. Integrated Ocean
Observing System (IOOS; https://ioos.noaa.gov/). Essential environmental
parameters are also organized, coordinated and acquired through global
initiatives like the Group on Earth Observations Biodiversity Observation
Network [GEO BON; @scholes2012], the Census of Marine Life [CoML; @coml2010]
and the Global Ocean Observing System (GOOS; http://www.goosocean.org).
However, environmental drivers, *i.e.* any externality that affects environmental processes
and disturbs natural systems and often referred to as stressors or pressures, remain
conspicuously missing from most of these initiatives [but see @halpern2015].
This is in spite of the need for high resolution data on their distribution and
intensity at a scale suitable for environmental management and impact assessments.

Drivers may be natural (e.g. sea surface
temperature anomalies and hypoxia) or anthropogenic (e.g. fisheries and marine
pollution) in origin.
The potential for complex interactions between drivers remains the largest
uncertainty when studying or predicting environmental change
[@darling2008; @cote2016]. The effects of multiple drivers can combine
non-linearly and result in effects that are greater (synergistic effect) or
lower (antagonistic effect) than the sum of individual effects
[@crain2008; @darling2008; @cote2016]. The uncertainty associated with complex
driver interactions must therefore be taken into account [@cote2016], yet most
research on driver effects in marine environments remains overwhelmingly focused
on single driver assessments [@obrien2019]. This is partly explained by the
complexity of studying multiple drivers simultaneously [@cote2016], but also
by the need for standardized data management practices and tools tailored to
the study of the effects of multiple drivers [@dafforn2016; @stock2018].

To address these issues, we are launching an initiative called *eDrivers*, an
open data platform focused on the description of the spatial distribution and
intensity of drivers and using the Estuary and Gulf of St. Lawrence in Canada
as a case study. The overarching goal of this platform is to create an expert
community committed to structuring, sharing and standardizing drivers data in
support to science and management.

The objective of this paper is twofold. The first is to present the guiding
principles underlying the inception and structure of *eDrivers*. We also wish
this paper to serve as a call for collaboration to any person, group or
organization who might be interested in contributing to this initiative. The
second objective focuses using the platform content to describe the intensity and distribution of drivers
in the St. Lawrence system and aims at 1) identifying areas of high cumulative exposure,
2) at characterizing areas with similar cumulative exposure regimes, dubbed threat
complexes [@bowler2019] and 3) at identifying drivers that are likely to interact
in the St. Lawrence.

# The St. Lawrence System

The St. Lawrence system, formed by one of the largest estuaries in the world and a
vast interior sea, is an complex social-ecological system characterized by highly
variable environmental conditions and oceanographic processes, both in space
and time [@el-sabh1990; @white1997; @dufour2007]. It thus offers a unique and
heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. It concomitantly provides
a wealth of ecosystem services that have historically and contemporarily
benefited the Canadian economy. It holds a rich fisheries industry targeting
over 50 species, serves as the gateway to eastern North-America by granting
access to more than 40 ports, has a booming tourism industry and an expanding
aquaculture production, develops emerging activities and has a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017]. With major
investments recently made and more forthcoming in economic and infrastructure
development and research [*e.g.* @goq2015; @rqm2018], an intensification of the
human footprint is expected in the St. Lawrence system.

Arising from this expected intensification and the experience of past and
current ecological tragedies [*e.g.* the collapse of cod fisheries and the
decline of the beluga population; @frank2005; @plourde2014; @dempsey2018] is a
call for an integrative and holistic approach to manage the St. Lawrence as a
whole. Yet, jurisdiction over the St. Lawrence is divided between the Canadian
federal government and five provincial governments (Newfoundland and
Labrador, New Brunswick, Nova Scotia, Prince Edward Island, and Québec) with
each governmental entity further divided into multiple departments.
Researchers from dozens of academic institutions and
countless local organizations also share an interest in the St. Lawrence. This
wide range of actors results in highly scattered data management
and sharing processes, which we wish to address through this open data platform
initiative.

# Open Data Platform

## Guiding principles

Five guiding principles motivate the development of the platform and shape its
structure (Figure \ref{diag}).

### Unity and inclusiveness

***Why***: Operating over such large scales in time, space and subject matter
requires a vast and diverse expertise that cannot possibly be possessed by any
one individual or organization. Consequently, we envision an initiative that
seeks to mobilize all individuals and entities with relevant expertise.

***How***: By promoting, consolidating and working with experts involved in
existing and highly valuable environmental initiatives already in place in the
St. Lawrence. Notable examples of environmental initiatives are the annual
review of physical [@galbraith2018], chemical, and biological [@blais2018]
oceanographic conditions in the St. Lawrence, the fisheries monitoring program
[@dfo2016], the annual groundfish and shrimp multidisciplinary survey
[@bourdages2018], the characterisation of benthic [@dutil2011], epipelagic and
coastal [@dutil2012] habitats of the St. Lawrence, and Canada's shoreline
classification [@eccc2018]. There are also nascent efforts to make available
information on several human activities in the St. Lawrence such as the Marine
Spatial Data Infrastructure portal, which provides data on zoning, shipping,
port activities, and other human activities in Canadian waters, including the
St. Lawrence system [@goc2016].

By working with existing data portals whose objective is to share
environmental data. We are thus collaborating actively with the St. Lawrence
Global Observatory (SLGO) to develop the initiative and to host the
platform on their web portal (https://ogsl.ca/en). The mission of SLGO is to
promote and facilitate the accessibility, dissemination and exchange of official
and quality data and information on the St. Lawrence ecosystem through the
networking of organisations and data holders to meet their needs and those of
users, to improve knowledge and to assist decision-making in areas such as
public safety, climate change, transportation, resources and biodiversity
conservation. The SLGO is also one of three regional association spearheading
the Canadian Integrated Ocean Observing System
(CIOOS; http://meopar.ca/research/cioos-call-for-proposals/),
which will focus on integrating oceanographic data from multiple sources to
make them accessible to end-users and to enable the national coordination of
ocean observing efforts by integrating isolated or inaccessible data, and by
identifying gaps or duplications in observations and research efforts.
We are also developing collaborations with the Portal on water knowledge
(http://www.environnement.gouv.qc.ca/eau/portail/), an initiative from the
Québec provincial government. This portal aims at collecting and sharing
accurate, complete and updated ressources on water and aquatic ecosystems to
support the mandate of relevant actors and stakeholders working in
water and aquatic ecosystems management.

By actively inviting, seeking, and developing collaborations as well as
encouraging constructive criticism from the inception and throughout the
lifetime of the platform.

By inviting external community contributions (Figure \ref{diag}).
External researchers or entities wishing to submit marine data will be able to
do so through the SLGO web portal (https://ogsl.ca/en). Submissions through
other data portals will also be accepted either through the development of data
sharing agreements or with the caveat that shared data are under an open source
license and that they adhere to the platform data standards.

### Findability, accessibility, interoperability and reusability

***Why***: Open data has been propelled to the forefront of scientific research
in an era of open, collaborative and reproducible science. By moving towards
large scale, cross-disciplinary research and management projects, there is a
growing need to increase the efficiency of data discovery, access,
interoperability and analysis, for both humans and computationally automated
processes [@reichman2011; @wilkinson2016]. Embracing open-data accelerates research
outcomes, improves transparency and replicability, assists in discovering
erroneous data and corroborating study findings, and decreases overall costs of
research initiatives [@reichman2011; @poisot2013; @lowndes2017]. As such, many
scientific journals (*e.g.* Frontiers, PLoS, Nature, Science), governmental agencies <!-- mettre nom de la revue ou on soumet -->
[*e.g.* @goc2016] and other institutions are adopting open
data-sharing policies and detailed management plans as part of their mandate.
Our goal is to foster efficient and functional open science by creating a
fully open, transparent and replicable open data platform.

***How***:
By building an infrastructure adhering to the FAIR Data Principles,
which states that data and metadata must be Findable, Accessible, Interoperable
and Reusable. These principles focus on the ability of humans and machines to
automatically find and (re)use data and knowledge [@wilkinson2016].

By making data and associated tools accessible through a variety of ways, namely
the SLGO web portal (https://ogsl.ca/en), two R packages called *eDrivers*
(https://github.com/orgs/eDrivers/eDrivers) and *eDriversEx*
(https://github.com/orgs/eDrivers/eDriversEx) to access the data through SLGO's
API and to provide analytical tools to explore data, respectively, and a Shiny
application (https://david-beauchesne.shinyapps.io/eDriversApp/) that allows
users to explore drivers data interactively (Figure \ref{diag}).
Note that the data are currently contained within and accessible through the
*eDrivers* R package only, as we are actively working to allow users to
download selected layers from SLGO's web portal and geoserver. The functions
available in *eDrivers* to access the data have however been developed to
ensure forward compatibility once the data migrate to SLGO's geoserver.

By defining clear data and metadata standards and specifications to support
the regional standardization of protocols and practices and to favour
interoperability with national and international programs like the SLGO web
portal, the upcoming CIOOS, Essential Ocean Variables (EOV) identified by the
Global Ocean Observing System [GOOS; http://www.goosocean.org] and the Ocean
Best Practices repository [OBP; https://www.oceanbestpractices.net/] from the
International Oceanographic Data and Information Exchange (IODE). The standard
used for the platform will thus be the North American Profile of ISO
19115:2003 - Geographic information - Metadata, a schema favoured for
geospatial data in Canada and the United-States.

By providing version control and code access to the workflows set up to generate
driver layers from raw data, the R packages and the Shiny application through
a GitHub organization called *eDrivers* (https://github.com/orgs/eDrivers/).

By committing to provide an increasingly FAIR infrastructure as we continue
developing it.

### Adaptiveness

***Why***: In the face of uncertainty and in an effort to address impending
environmental changes, adaptive management has been identified as the chief
strategy to guide efficient decision-making
[*e.g.* @costanza1998; @margules2000; @keith2011; @jones2016] and has already
been discussed in the context of multi-drivers and cumulative impact assessments
[@halpern2015; @beauchesne2016; @cote2016; @schloss2017]. Adaptive management can only be
truly achieved through a commitment to adaptive monitoring and data reporting
[@margules2000; @halpern2012; @lubchenco2015]. We further contend that adaptive
management requires the development of adaptive monitoring tools and
infrastructures, which we seek to address through a continuously-evolving platform.

***How***:
By setting up mechanisms structuring cyclic reviews of platform content,
for the integration of new material (*e.g.* data and methods) as it becomes
available or accessible, and by striving to provide time-series data that are
crucial to assess temporal trends and potentially early-warning signals of
ecosystem change (Figure \ref{diag}).

### Validation

***Why***: Using high-quality data is crucial to ensure robust analyses and
decision-making [@cai2015; @veiga2017]. Data quality assessment (*i.e.* quality
measurement, validation and classification) and management (*i.e.* quality
control and assurance) are therefore crucial steps in data sharing [@veiga2017].

***How***: By selecting a framework to assess and manage data quality and by
providing a data quality report with for all driver layers shared on the
platform.

### Recognition

***Why***: Like peer-reviewed publications, data must also be given its due
importance in scientific endeavors and thus be considered as legitimate citable
products contributing to the overall scientific output of data providers
[@codata2013; @force112014]. Appropriate citations should
therefore be provided for all data layers used and shared by the platform.

***How***: By including the persistent identifier, when possible, and full
citation of all data layers used and shared on the platform.

By listing all known projects and peer-reviewed publications using or citing
the data shared on the platform.

By adhering to the Data Citation Principles [@force112014], which
focus on citation practices that cover purpose, function and attributes of
citations in an effort to make citations human and machine readable as well as
providing appropriate credit to data products. As such, all data layers shared
on the platform will be identified by a persistent identifier (*e.g.* DOI),
possess a license outlining conditions of use and a clear *how to cite* label
including the authors, the data version and year, an informative title, the
persistent identifier and how to access the data (*e.g.* url).

# St. Lawrence drivers

## Data

The list of drivers for which we sought data was informed by global cumulative
impacts assessments initiatives [@halpern2008; @halpern2015], regional holistic
evaluation of the state of the St. Lawrence [@dufour2007; @benoit2012], and
through communications with regional experts. Drivers are varied in origin,
from terrestrial (*e.g.* nutrient input) to marine (*e.g.* shipping), and from
anthropogenic (*e.g.* fisheries) to natural (*e.g.* temperature anomalies).
Through the data gathering process, we developed and continue developing
collaborations with regional experts and data holders.
We also use global data from seminal works on cumulative impacts on global
habitats [@halpern2008; @halpern2015] and available on the National Center for
Ecological Analysis and Synthesis (NCEAS) online data repository
[Table 1; @halpern2015b]. We selected global data that
were unavailable at the regional scale and that were available at a resolution
adequate for use at the scale of the St. Lawrence (*e.g.* marine pollution).

The platform currently provides access to 22 driver layers (Table 1) and we are
actively working on updating or developing additional driver layers.
We divided drivers into 4 groups: coastal, climate, fisheries and marine traffic
related drivers (Table 1). All data layers and methodologies used to produce
driver layers are described in the supplementary materials of this paper and on
the *eDrivers* GitHub organization (https://github.com/orgs/eDrivers/).
Also note that the specific
methodologies underlying the figures presented in the paper are described in
the figure captions to ease the flow of text.

## Cumulative exposure

Apart from the northeastern Gulf, cumulative exposure (*i.e.* sum of normalized
driver intensity) is ubiquitous in the St. Lawrence (Figure \ref{footprint}).
Cumulative exposure is generally highest along the coast (Figure \ref{footprint}),
with most intense hotposts (*i.e.* areas of overlapping drivers at high relative
intensity) located in the vicinity of coastal cities (Figure \ref{hotspot}).
In general, offshore areas are less exposed to cumulative drivers, with the
Estuary and the Anticosti Gyre being notable exceptions (Figures \ref{footprint}
and \ref{hotspot}). That is however not to say that offshore areas are free of
exposure, as most of the St. Lawrence is exposed to multiple drivers
(Figures \ref{footprint} and \ref{hotspot}).

These results are consistent with observations elsewhere in the world, where
cumulative exposure notably arises from and markedly intensifies close to
coastal cities and at the mouth of rivers draining highly populated areas
[e.g. @halpern2015; @feist2016; @mach2017; @stock2018]. These are areas
where human uses (*e.g.* coastal development and shipping) and footprint
(pollution runoff) are the most intense [@feist2016], and on which is
overlaid a background of natural disturbances [@micheli2016]. They are also the
areas in which the most dramatic increases in exposure are expected, with
populations increasing more rapidly along the coast than inland [@feist2016].

In the St. Lawrence, large coastal cities are mostly located along the Estuary
and the southwestern Gulf, while the northeastern Gulf is largely uninhabited
or home to small coastal towns. Moreover, the Estuary, along with the
St. Lawrence river, provide access to and serve as the primary drainage
outflow of the Great Lakes Basin, which is the most densely populated region
in Canada [@statcan2017].

## Driver interactions

By providing data on driver intensity and distribution, the platform will
greatly facilitate the study of the effects of multiple drivers, their interaction
likelihood and the areas where they may interact.
For instance, drivers like hypoxia (Figure \ref{kernel}A) and demersal destructive
fisheries (Figure \ref{kernel}B) are known to occur mainly in deeper areas of the
St. Lawrence, and hence an interaction between the effects of the two drivers
could be anticipated. Fisheries in the St. Lawrence have historically affected
biodiversity distribution and habitat quality [@moritz2015]. Hypoxia, meanwhile,
decreases overall habitat quality but has variable effects that are
species-dependent, ranging from well-adapted [*e.g.* northern shrimp
*Pandalus borealis* and Greenland halibut *Reinhardtius hippoglossoides*;
@pillet2016] to reduced growth rates [@dupont2013] and avoidance of oxygen
depleted habitats [*e.g.* Atlantic cod *Gadus morhua*; @chabot2008] to increased
mortality [*e.g.* sessile benthic invertebrates; @eby2005; @gilbert2007; @belley2010].
Certain species may thus be negatively affected by fisheries and withstand hypoxia
but still experience a decrease in prey availability, while others may be
negatively affected by the compounded effect of both drivers [@deleo2017].

By combining both drivers (Figure \ref{kernel}C), we can evaluate
which regions in the St. Lawrence are more likely to be affected by their
simultaneous presence and evaluate the relative intensity at which they tend to
co-occur (Figure \ref{kernel}D). Doing this, we observe that hypoxia and bottom
fisheries tend to co-occur in the St. Lawrence and that they do so at relatively
high intensities, with extensively used fishing grounds also characterized by
low levels of oxygen neighboring the limits of hypoxia (*i.e.* ~0.7 mg*L^-1^,
corresponding to ~30% saturation). We demonstrate the ease with which this
can be accomplished using *eDrivers* by providing the code necessary to reproduce
Figure \ref{kernel} in box 1.

## Threat complexes

The number of driver overlapping in the St. Lawrence increases with cumulative
exposure (Figure SX). Areas with high exposure such as the Estuary, the
Anticosti Gyre and the southwestern Gulf coastline (Figure \ref{footprint} and
\ref{hotspot}) are thus areas in the St. Lawrence where driver interactions are
most likely, and where they can arise between a host of different drivers.
While we cannot ascertain that high exposure areas are also the most impacted,
we can safely predict that these are the areas where studying ecosystem state
will be the most complex due to the uncertainty associated with driver
interactions, an uncertainty certain to increase rapidly with the increasing
number of interacting drivers [@cote2016]. It is however important to note
that there are very few areas of the St. Lawrence that do not experience driver
overlap (Figure \ref{hotspot}) and that drivers do not necessarily need to be
co-occurring at high intensities to cause significant adverse effects
[*e.g.* @relyea2009; @liess2016].

In an effort to describe areas where interactions are most likely between
multiple and typical combinations of drivers, we identified areas with similar cumulative
exposure regimes, which we refer to as threat complexes after @bowler2019. Seven
distinct threat complexes were identified in the St. Lawrence, which can be
divided into 4 offshore and 3 coastal complexes (Figure \ref{cluster}). Refer
to the supplementary information for the figures related to the following
description of threat complexes.

Coastal threat complexes (2, 4 and 6; Figure \ref{cluster}) are differentiated
from offshore complexes by the inclusion of
all coastal drivers. They however also include all other types of drivers except
hypoxia. They are also the most exposed threat complexes, both in terms of
driver overlap and intensity. Among coastal complexes, complex 2 is
differentiated by the presence of aquaculture sites. Threat complex 4, which is
widespread through the coastal Estuary and Gulf, includes areas associated with
port activities at higher relative intensity. Finally, complex 6, in the
southeastern and northernmost Gulf, is characterized by negative bottom
temperature anomalies.

Offshore threat complexes (1, 3, 5 and 7; Figure \ref{cluster}) are generally
characterized by climate and marine traffic related drivers. Threat complex 1
mostly incorporates the shallow areas of the northern Gulf and is characterized
by lower marine traffic intensity and with higher intensity surface temperature
anomalies. Threat complex 3 encompasses most of the southern Gulf and is
characterized by medium to high marine traffic intensity and higher aragonite
stress. Threat complex 5, meanwhile, is mostly delineated by the deep channels
of the northern Gulf and is characterized by positive bottom temperature
anomalies. Finally, complex 7 corresponds to the Laurentian channel, which is a
deep channel connecting the Estuary to the Atlantic. This threat complex is
characterized by high intensity hypoxia, marine traffic and pollution, and
demersal destructive fisheries.

Fisheries are generally distributed among all offshore threat complexes and
coastal complex 4. This widespread distribution among clusters causes fisheries
to have limited contribution to differences between complexes. There are however
certain threat complexes characterized by higher proportion of fishing
activities. Threat complex 3 holds higher intensities of pelagic low-bycatch and
demersal high-bycatch fisheries. Threat complex 5, meanwhile, has higher intensities of
demersal destructive fisheries. Finally, threat complex 7 is characterized by higher intensities
of demersal destructure and pelagic high-bycatch fisheries.

Describing these threat complexes allows us to see that each of those areas
are characterized by distinct overlapping suites of drivers. This holds
significant importance for the species, habitats and ecosystems located
within or moving through those threat complexes, as they are likely to be
affected by the joint effects of all drivers typical to those
complexes [@bowler2019].

# Perspectives and next steps

Understanding how ecosystem state will be affected by global change necessitates
a comprehensive understanding of how threats are distributed and interacting in
space and time, which hinges on appropriate data tailored to multi-drivers
studies [@dafforn2016; @stock2018; @bowler2019]. In the St. Lawrence,
we found that few areas are free of cumulative exposure and that the whole
Estuary, the Anticosti Gyre and coastal southwestern Gulf are particularly
exposed to cumulative drivers, especially close to urban areas. We also
identified seven geographically distinct threat complexes that show similar
cumulative exposure regimes. These complexes also show that coastal areas are
particularly exposed to all types of drivers and that multiple drivers are
typically co-occurring in space. The *eDrivers* initiative allows us to
efficiently identify these areas in need of heightened scrutiny. It also ensures
an iterative improvement of cumulative exposure assessments and threat complexes
identification as gaps in knowledge are addressed and approaches to describe
drivers are refined.

Arguably the most meaningful benefit anticipated from *eDrivers* is the
significant gain in data accessibility through the creation of a
cross-jurisdictional collaborative tool. This will effectively curtail the need
to reach dozens of experts across multiple organizations and over extensive
periods of time to assemble the data needed for systematic research. These
benefits could pay quick scientific and management dividends by drawing on the
knowledge and efforts of a wide range of contributors, by expanding avenues of
scientific inquiry, by decreasing overall effort duplication and research costs,
and by increasing research efficiency [@franzoni2014]. Critically - and we
emphasize this point - the platform will allow the scientific and governmental
communities to identify key knowledge gaps that will assist in prioritizing
and optimizing research efforts. For example, the platform could assist in
designing targeted experiments to study the effects of multiple drivers by
identifying sampling sites capturing appropriate exposure gradients
[@stock2018; @bowler2019].

Another key issue with data sharing that will be addressed by the platform is
data interoperability. By using and enforcing (meta)data standards the platform
will strongly encourage the standardization of future data collection protocols
to maintain compatibility. Standardized protocols will ease the regional
aggregation of local datasets that are currently highly fragmented in
availability, quality and experimental approaches [@reichman2011]. It will also
enable interoperability with other open data platforms (*e.g.* OBIS). This
standardization could also be accompanied by an initiative to extend
the Essential Ocean Variables (http://www.goosocean.org) to include a list of
essential driver variables in marine environments. Additional steps towards
standardization could also be taken by emulating protocols elaborated through
international programs like GEO BON [@scholes2012] and CoML [@coml2010].

We also believe that the *eDrivers* initiative will operationalize
evidence-based decision-making by streamlining data management and research,
allowing science output to be available and interpretable on a time scale
relevant to management. The platform will thus greatly facilitate the
application of broad scale, holistic research and management approaches such as
ecosystem-based management, strategic environmental assessments and
social-ecological metanetworks [*e.g.* @halpern2015; @jones2016; @dee2017].

Significant effort is still needed to bring our vision to fruition. Foremost on
that list is to maintain our efforts to foster collaborations, develop platform
content and identify key knowledge gaps. A fair and efficient organizational
structure will also be developed in order to manage the platform as a community.
Appropriate funding must also be secured to continue building this community and
ensure the long-term viability of the initiative, although the partnership with
SLGO partly addresses this issue.

Finally, terrestrial and coastal environments must also be incorporated, as
sources of stress within those habitats extend to the marine environments.
Moreover, despite coastal areas being recognized as the most exposed to
environmental threats we continue delineating
terrestrial and marine realms, using coastlines as we would an impermeable
barrier. While there is a sensible rationale for this division, we must strive
to eliminate this barrier if we are to appropriately study and predict the
impacts of global change [*e.g.* see @bowler2019].

Despite the work still ahead, we are hopeful that this initiative will be a very
successful one. Ultimately, this platform represents a much needed solution to
address important issues in data management that could radically shift broad
scale research and management practices towards efficient, adaptive and holistic
ecosystem-based management in the St. Lawrence and elsewhere in the world. All
it requires to be successful is for the scientific and political communities to
fully commit to open data, adaptive monitoring and, most of all, an integrated
vision of ecosystem management.

# Acknowledgements

We thank the Fond de Recherche Québécois Nature et Technologie (FRQNT) and the
Natural Science and Engineering Council of Canada (CRSNG) for financial support.
This project is also supported by Québec Océan, the Quebec Centre for
Biodiversity Science (QCBS), Takuvik, and the Notre Golfe networks. This
research is also sponsored by the NSERC Canadian Healthy Oceans Network and its
Partners: Department of Fisheries and Oceans Canada and INREST (representing the
Port of Sept-Îles and City of Sept-Îles).

# Author contributions statement

TO WRITE

# Conflict of interest statement

The authors declare that the submitted work was carried out in the absence of any personal, professional or financial relationships that could potentially be construed as a conflict of interest.

\newpage

# Listings

Box 1. Code snippet demonstrating how to use the *eDrivers* to reproduce
figure \ref{kernel} in R.

\begin{lstlisting}[frame=single]
# Install and load eDrivers package
devtools::install_github('eDrivers/eDrivers')
library(eDrivers)

# Load data
drivers <- fetchDrivers(drivers = c('hypoxia','fishDD'))

# Get data from `eDrivers` class object
driverData <- getData(drivers)

# Normalize data
driverData <- driverData / cellStats(driverData, 'max')

# Visualize data and combination
plot(driverData$fishDD)   # Demersal destructive fisheries
plot(driverData$hypoxia)  # Hypoxia
plot(sum(driverData))     # Combination

# Identify values > 0 and not NAs
driverData$fishDD[driverData$fishDD < 0] <- NA
driverData$fishDNH[driverData$hypoxia < 0] <- NA
id0 <- !is.na(values(driverData$fishDD)) &
       !is.na(values(driverData$hypoxia))

# 2D kernel for driver co-intensity
library(MASS)
coInt  <- kde2d(x = values(driverData$fishDD)[id0],
                y = values(driverData$hypoxia)[id0],
                n = 500, lims = c(0, 1, 0, 1))
image(coInt, zlim = c(0,max(coInt$z)))

# Driver density distribution
plot(density(driverData$fishDD[id0]))   # Demersal destructive
plot(density(driverData$hypoxia[id0]))  # Hypoxia
\end{lstlisting}

\newpage

# Figures

![Diagram of the platform structure. Community input in the form of raw data is accessed through the St. Lawrence Global Observatory (SLGO); https://ogsl.ca/en) repository - the platform host - or through open access repositories (*e.g.* NASA data). The raw data are then processed through a workflow hosted on the *eDrivers* GitHub organization (https://github.com/orgs/eDrivers/). Data processing may be as simple as data rescaling (*e.g.* night lights) or make use of more complex methodologies (*e.g.* waste water). All data is then hosted on SLGO's geoserver and accessible through their API. We developed a R package called *eDrivers* to access the driver layers through R and we are actively developing a second R package called *eDriversEx* that includes analytical tools to explore drivers data. Finally, we have developed a Shiny app that allows users to explore drivers data interactively (https://david-beauchesne.shinyapps.io/eDriversApp/). All R components of the project are also hosted on the *eDrivers* GitHub organization. \label{diag}](./figures/structure.png)

\newpage

![Distribution of driver footprint in the Estuary and Gulf of St. Lawrence. The footprint is measured by summing the relative intensity of all drivers for each grid cell: $F_d = \sum_{i=1}^{n} D_i$. Driver layers were also log-transformed when their frequency distribution was non-normal. All driver layers were also normalized between 0 and 1 using the 99th quantile to allow for direct comparison of relative intensities between drivers and control for extreme values. \label{footprint}](./figures/footprint.png)

\newpage

![Distribution of hotspots of driver footprint in the Estuary and Gulf od St. Lawrence. Hotspots were identified using the sum of all driver's 80th intensity quantile. This results in the distribution of areas with overlapping drivers at high intensity. \label{hotspot}](./figures/hotspot.png)

\newpage

![Example of platform content using the spatial distribution and the intensity of hypoxia and demersal destructive fisheries in the St. Lawrence. An index of hypoxia (**A**) was created using bottom oxygen concentrations from a newly instated survey during the months of August and September [@starr2019]. We interpolated the survey data using cokriging with depth as a covariable [@dutil2011]. A hypoxic index was then measured using a hypoxic threshold of 2 $mlL^{-1}$ bottom oxygen concentration [@diaz1995]. An inverted logistic curve was used to create an index of hypoxic stress ranging from close to hypoxic (*i.e.* 0) to anoxic (*i.e.* 1), under the assumption that hypoxic stress increases following a logistic curve as it approaches the hypoxic threshold. Demersal destructive fisheries (**B**) correspond to fishing activities using trawl and dredges. Intensity of fishing activities was measured using fisheries location data collected between 2010 and 2015 [@dfo2016] and a 2km buffer to consider trawling and dredging area. For each study area cell ($1 km^2$), fishing intensity (FI) was measured as the annual area weighted total biomass ($kg$) yield for all species captured: $FI = \sum_{i = 1}^{n} B_i * \frac{Area_{Fish}}{Area_{Cell}}$, where $B_i$ is the total annual biomass captured in each cell. As fishing intensity was highly skewed towards zeroes, we FI was log transformed to obtain a workable distribution. Values were then normalized to allow comparison with hypoxia values. Drivers intensity was summed (**C**) to visualize the combination of hypoxia and demersal destructive fisheries. Finally, individual density and the co-intensity of hypoxia and demersal destructive fisheries was investigated with a kernel analysis (**D**).\label{kernel}](./figures/kernel.png)

\newpage

![Distribution of threat complexes in the Estuary and Gulf of St. Lawrence. Threat complexes (a term coined by [@bowler2019]) are areas with similar cumulative driver exposure regimes. Threat complexes were identified using a partional *k-medoids* clustering algorithm, CLARA [CLustering for Large Applications; @kaufman1990]. The appropriate number of clusters $k$ was tested using a range of values and validated by selecting the number of clusters that maximized the average silhouette width [@kaufman1990] and minimized the total within-cluster sum of squares (WSS). We also validated the clustering by comparing *k-medoids* clustering with *k-means* clustering with the *Lloyd* algorithm [@lloyd1982] and found a 66% cell correspondence between both approaches, with most of the differences attributable to a single undivided cluster in the *k-means* approach. Seven distinct threat complexes were identified in the St. Lawrence. The difference between clusters was explored by measuring the total inter-cluster dissimilarity and the contribution of each driver to the total inter-cluster dissimilarity through a similarity percentage analysis (SIMPER) with Bray-Curtis dissimilarity [@clarke1993]. As the drivers data is too large, we used a bootstrap procedure for the SIMPER analysis, randomly selecting 5% of each cluster to run the analysis and repeating the process over 300 iterations. A similar bootstrapping procedure was used to evaluate the intra-cluster similarity with the Bray-Curtis similarity using 300 iterations using 5% of each cluster data when the total number of observations in cluster exceeded 10000 observations. The *k-medoid*, *k-means* and SIMPER analyses were performed using the *cluster*, *stats* and *vegan* R packages, respectively [@maechler2018; @oksanen2018; @rcoreteam2018]. Refer to the Supplementary Materials for more details. \label{cluster}](./figures/cluster.png)

\newpage
\blandscape

# Tables

Table 1. List of drivers currently available on *eDrivers* and used for the analyses presented in this paper.

```{r table1, echo=F, fontsize=4}
load('./tables/table1.RData')
table1
```

\elandscape
\newpage

# References

<!--
Additional notes:

*Elements needed*:
  - Platform host (F-A-I-R): SLGO, (present CIOOS)
X  - Data access (F-A-I-R):
X    - SLGO: Rawdata, Geoserver, API
X    - eDrivers R package
X    - eDriversEx R package
X    - Shiny application
X  - Code access and version control (F-A-I-R):  eDrivers organization
X    - Drivers workflow, R packages, Shiny app
O  - Essential Drivers Variables (I): EOV, GOOS (EBV)
B1  - Rich metadata standard and ontology (F-I-R): SLGO, CIOSS, Ocean Best Practices(?)
B1  - Data and metadata format (I-R): JSON & TXT
B1  - Data citation (F-R): Data citation principles Force11
B1  - Persistent Identifier (F): DOI for all datasets used (if available) and shared
B1  - License agreement (A-R): detailed for each data shared
B1  - How to cite

Reusability:
- Describe the scope of your data: for what purpose was it generated/collected?
- Mention any particularities or limitations about the data that other users should be aware of.
- Specify the date of generation/collection of the data, the lab conditions, who prepared the data, the parameter settings, the name and version of the software used.
- Is it raw or processed data?
- Ensure that all variable names are explained or self-explanatory (i.e., defined in the research field’s controlled vocabulary).
- Clearly specify and document the version of the archived and/or reused data.
- Where the data came from (i.e., clear story of origin/history, see R1),
- who to cite and/or how you wish to be acknowledged.
- Include a description of the workflow that led to your data:
  - Who generated or collected it? How has it been processed?
  - Has it been published before?
  - Does it contain data from someone else that you may have transformed or completed?
  - Ideally, this workflow is described in a machine-readable format.
 -->