### 2019-03-29 eDriversMS-v2.Rmd
add updated introduction

---
title: Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^5^, Simon Bélanger^4^, Pascal Bernatchez^4^, Hugo Bourdages^7^, Clément Chion^6^, Peter S. Galbraith^7^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^7^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^7^, Anne-Sophie Ste-Marie^5^, Philippe Archambault^2^
date: 2019-02-06
fontsize: 12pt
output:
  pdf_document:
    toc: false
    number_sections: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   # - \usepackage{titling}
   # - \pretitle{\begin{flushleft}}
   # - \posttitle{\end{flushleft}}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
relativeurls: true
bibliography: eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

<!--
cd phdproj/eDrivers/eDriversMS/
rmarkdown::render('./eDriversMS.rmd', 'pdf_document')
rmarkdown::render('./eDriversMS.rmd', 'word_document')
-->

<!-- Setting up parameters for code snippet presented in Box 1 -->
\lstset{language=R,
    basicstyle=\small\ttfamily,
    stringstyle=\color{DarkGreen},
    otherkeywords={fetchDrivers, install_github, getData, kde2d},
    morekeywords={TRUE,FALSE},
    deletekeywords={data,frame,length,as,character},
  keywordstyle=\color{blue},
  commentstyle=\color{DarkGreen},
}

^1^Institut des sciences de la mer, Université du Québec à Rimouski, Rimouski, QC, Canada \newline
^2^Québec Océan, Département de biologie, Université Laval, Québec, QC, Canada \newline
^3^Département de biologie, Université de Sherbrooke, Sherbrooke, QC, Canada \newline
^4^Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \newline
^5^St. Lawrence Global Observatory, Rimouski, QC, Canada \newline
^6^Département des Sciences naturelles, Université du Québec en Outaouais, Gatineau, QC, Canada \newline
^7^Fisheries and Oceans Canada, Maurice Lamontagne Institute, Mont-Joli, QC, Canada \newline
^8^National Center for Ecological Analysis and Synthesis, University of California, Santa Barbara, CA, United States \newline
^9^Bren School of Environmental Science and Management, University of California, Santa Barbara, CA, United States \newline
^10^Department of Earth & Planetary Sciences, McGill University, Montréal, QC, Canada \newline
^11^Ministère Environnement et Lutte contre les changements climatiques, Québec, QC, Canada  \newline

**Correspondence**: \newline
David Beauchesne \newline
david.beauchesne@uqar.com \newline

**Number of figures**: 5 \newline
**Number of tables**: 1 \newline

\newpage

# Abstract

# Introduction

The St. Lawrence System, formed by one of the largest estuaries in the world and
a vast interior sea, is a complex social-ecological system characterized by
highly variable environmental conditions and oceanographic processes, both in
space and time [@el-sabh1990; @white1997; @dufour2007]. It thus offers a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
Lawrence System provides a wealth of ecosystem services that have historically
and contemporarily benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017 @archambault2017]. With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @goq2015; @rqm2018], an
intensification of the human footprint is expected in the St. Lawrence System.

As elsewhere in the world [see @halpern2015], this intensifying human footprint
will likely result in increasingly intricate environmental exposure regimes,
*i.e.* suites of overlapping and potentially interacting environmental drivers
threatening ecosystems, habitats or ecological communities. Drivers, often
referred to as stressors or pressures, are any externalities that affect
environmental processes and disturb natural systems. Drivers may originate from
natural or human-induced biophysical processes (*e.g.* sea surface temperature
anomalies and hypoxia) or directly from anthropogenic activities (*e.g.*
fisheries and marine pollution). The potential for complex interactions between
drivers is the largest uncertainty when studying or predicting environmental
change [@darling2008; @cote2016]. The effects of multiple drivers can combine
non-linearly and result in effects that are greater (synergistic effect) or
lower (antagonistic effect) than the sum of individual effects
[@crain2008; @darling2008; @cote2016].  <!-- For example...??? -->

The uncertainty associated with complex driver interactions must therefore be
taken into account when investigating environmental impacts [@cote2016], yet
most research on driver effects in marine environments remains overwhelmingly
focused on single driver assessments [@obrien2019]. Arising from the expected
increase in environmental exposure and the experiences of past and
current ecological tragedies such as the collapse of cod fisheries
[@frank2005; @dempsey2018] and the decline of the beluga and right whale
populations [@plourde2014] is a need to characterize the distribution,
intensity and overlap between drivers in the St. Lawrence System. This will
provide critical information on areas most exposed to cumulative drivers and on
the interaction potential of drivers in the St. Lawrence. It is also a necessary
step towards the holistic and integrated management of the St. Lawrence System.

Gathering environmental data for large scale, systematic initiatives can,
however, be a very challenging and time consuming - not to say painful - process.
On one hand, there is an overwhelming and expanding wealth of data available.
Such information overload may inhibit our ability to make decisions based on
scientific information, promote massive effort duplication, disproportionately
appropriate research funds to certain sectors, and obscure knowledge gaps amid
a sea of information [@eppler2004]. On the other hand, crucial data are lacking
and remain largely unavailable or inaccessible for a variety of reasons,
including proprietary rights, lack of organizational time, capacity and
training, and in some rare cases unwillingness to share, curtailing our ability
for appropriate decision-making.

The current digital infrastructure is highly decentralized and the data
management and sharing practices are highly heterogeneous, preventing us from
maximizing benefits from research investments [@wilkinson2016].
Yet there now exists multiple initiatives that address this issue by assembling,
organizing and sharing vast arrays of environmental data.
Biotic data can be accessed through web portals such as the Ocean Biogeographic
Information System [OBIS; @obis2018], the Global Biodiversity Information Facility
[GBIF; @gbif2018], the Global Biotic Interactions platform [GloBI; @poelen2014],
and the World Register of Marine Species [WoRMS; @worms2017]. Abiotic data can
also be accessed through WorldClim [@hijmans2005], Bio-ORACLE
[@tyberghein2012], and MARSPEC [@sbrocco2013]. Initiatives focused on sharing
environmental data for specific areas also exist, such as the St. Lawrence
Global Observatory (SLGO; https://ogsl.ca/en), the European Marine Observation
and Data Network (EMODnet; http://www.emodnet.eu/) and the U.S. Integrated Ocean
Observing System (IOOS; https://ioos.noaa.gov/). Essential environmental
parameters are also organized, coordinated and acquired through global
initiatives like the Group on Earth Observations Biodiversity Observation
Network [GEO BON; @scholes2012], the Census of Marine Life [CoML; @coml2010]
and the Global Ocean Observing System (GOOS; http://www.goosocean.org).

Tools collating data and knowledge on multiple drivers in a comparable and
interoperable way are however conspicuously missing from these initiatives
[but see @halpern2015]. This is in spite of integerated management and assessment
approaches needing efficient data reporting, standardized data management
practices and tools tailored to the study of the effects of multiple drivers
[@dafforn2016; @stock2018]. An additional objective thus emerged in the process of
addressing our initial goals:
sharing the knowledge gathered through the process of characterizing drivers in the St. Lawrence.
We are launching *eDrivers*,
an open knowledge platform focused on sharing knowledge on the distribution and
intensity of drivers in the St. Lawrence System and gathering a community of
experts committed to structuring, standardizing and sharing knowledge on drivers
in support of science and management.

This manuscript thus serves to present two distinct but related goals. The first
is to characterize the distribution and intensity of drivers in the St. Lawrence
System in order to: 1) identify areas of high cumulative exposure,
2) characterize areas with similar cumulative exposure regimes, dubbed threat
complexes [@bowler2019] and 3) identify drivers that are likely to interact
in the St. Lawrence. The second is to present the guiding
principles underlying the inception and structure of *eDrivers* and use
this paper as a call for collaboration to any person, group or
organization that might be interested in contributing to this initiative.