### 2020-04-24 Revisions2/Beauchesne-eDriversMS-SI.rmd
finish answer to reviewer comments

---
title: Supplementary Material
fontsize: 12pt
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  html_document:
    number_sections: true
header-includes:
    - \usepackage{float}
    - \usepackage{setspace}
    - \onehalfspacing
    - \usepackage[svgnames]{xcolor}
    - \usepackage[normalem]{ulem}
    - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

<!--
rmarkdown::render('./Revisions2/Beauchesne-eDriversMS-SI.rmd')
-->

# Drivers description

Table S1. List of drivers currently available on *eDrivers* along with their respective acronym used in the figures of the supplementary material.

```{r tableS1, echo=F, fontsize=4}
load('../tables/tableS1.RData')
tableS1
```

## Climate

### Acidification

Oceans are the largest reservoirs and sinks of atmospheric carbon dioxide ($CO_2$). Its uptake increases seawater acidity and lowers the saturation state of waters with respect to calcite ($\Omega_C$) and aragonite ($\Omega_A$), the two most common $CaCO_3$ polymorphs that constitute the shells and skeletons of many marine organisms [@mucci2017]. Ocean acidification can have deleterious effects on carbonate-secreting organisms (*e.g.* mollusks and crustaceans) and certain physiological processes in non-calcifying organisms [@fabry2008; @kroeker2013].

When $\Omega_C$ or $\Omega_A$ decrease below 1, water becomes undersaturated and corrosive to the skeletal minerals of carbonate-secreting organisms. Note that if $\Omega_A$ $< 1$, the waters may still be supersaturated with respect to calcite since it is 50% more soluble than aragonite. However, organisms have wide ranging responses to changes in $\Omega_C$ and $\Omega_A$. For example, most corals stop calcifying at $\Omega_A$ $< 2$ [@kleypas2006; @bove2019], while other organisms may have adapted to precipitate $CaCO_3$ even when seawater is undersaturated with respect to calcite [$\Omega_C$ $< 1$ or $\Omega_A$ $< 0.65$; e.g. @uthicke2016].

We used $\Omega_{A}$ to characterize ocean acidification in the bottom waters of the St. Lawrence. The data come from the Department of Fisheries and Oceans' (DFO) Atlantic Zone Monitoring Program [AZMP; @galbraith2018] surveys and DFO's multispecies surveys [@bourdages2018] collected in August-September of 2017 [@starr2019]. $\Omega_{A}$, pH, and dissolved oxygen ($O_2$) were measured at 117 stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The majority of sampling stations were located in the Southern Gulf ($n = 84$), the Estuary and Northern Gulf ($n = 33$).

The carbonate speciation was determined through pH and total alkalinity (TA) measurements. Samples for pH and TA were collected with no head space in 500 $mL$ borosilicate glass flasks, and 250 $\mu L$ of saturated $HgCl_2$ solution was added to each sample and processed following the "Guide to best practices for Ocean CO2 Measurements" [@dickson2007].

$pH_T$ was determined spectrophotometrically using the indicator dye m-cresol
purple (Sigma-Aldrich). Absorbance was measured at 730, 578 and 434 nm before
and after dye addition in 10 cm quartz cells thermostated at 25 ± $0.05^oC$
[@dickson2007]. A similar procedure was carried out before each set of sample
measurement using a TRIS (Tris (hydroxymethyl) -aminomethane) buffer prepared
at a practical salinity (S) of approximately 30 [@millero1986]. Certified
Reference Material (CRM) (supplied by Professor Andrew Dickson, Scripps
Institution of Oceanography, San Diego, USA) was used for quality control of
our pH TRIS buffer.

TA was determined by potentiometric titration in an open cell using an
automated Radiometer potentiometric titrator (Titrilab 865) and a pH combination
electrode (pHC2001) in a continuous titrant addition mode, an algorithm
specifically designed for shallow end-point detection [@dickson2007]. The
0.1M HCl titrant in a solution of 0.6M NaCl was calibrated using CRM provided by Professor Andrew G. Dickson.

The carbonate system parameters (including $\Omega_{A}$) were
corrected for *in situ* pressure and temperature using the algorithm CO2SYS
[@lewis1998] with measured pH, TA, soluble reactive phosphate and silicate concentrations as input parameters.

To account for the low sample size in the Northern Gulf, we used the correlation
between $O_2$ and $CO_2$, which are linked through the stoichiometry of the
respiration reaction in waters that are isolated from the atmosphere [@mucci2011].
Metabolic $CO_2$ increases in water through biotic processes such microbial
respiration of organic matter [@mucci2011]. In deep waters, found in the Laurentian, Anticosti and Esquiman Channels, variations of
$\Omega_{A}$ are largely governed by the addition of metabolic $CO_2$, whereas, near the air-sea interface, $O_2$ is replenished faster than $CO_2$ can escape
[@zeebe2001]. The impact of respiration on carbonate chemistry is highest in
hypoxic regions, where metabolic $CO_2$ accumulates to high concentrations
[@mucci2011].

We explored the correlation between $\Omega_{A}$ and $O_2$ in the data and found
no significant correlation at the scale of the whole St. Lawrence (*p-value*:
0.08; $R^2$: 0.03). Nevertheless, a correlation between $\Omega_A$ and $O_2$
exists in the Northern St. Lawrence (*p-value*: < 0.01; $R^2$: 0.49), but not
in the Southern St. Lawrence (*p-value*: 0.85; $R^2$: < 0.01). Furthermore, the
correlation observed in the north between $\Omega_A$ and $O_2$ is especially
strong when deep Channel waters (*Deep* model: *p-value*: < 0.01; $O_2$: 0.20; $R^2$:
0.97) are separated from shallower regions (*Shallow* model: *p-value*: < 0.01;
$O_2$: 0.07; $R^2$: 0.96), with $\Omega_A$ increasing more steeply with
increasing $O_2$ in the deep Channels. This stronger correlation is likely caused by the lack of gas exchange with the atmosphere or the surface mixed layer and the importance of metabolic $CO_2$ in deep waters, some of which may have mean ages of up to 20 years [@mucci2011]. In the Northern Gulf, the acidification trends with depth and distance from the Cabot Strait along the Laurentian, Anticosti and Esquiman Channels due to the progressive oxygen depletion and metabolic carbon dioxide accumulation [@mucci2011].

There are likely different dynamics at play in the Southern Gulf, such as waters with lowered alkalinity and salinity caused by the surface, seaward-flowing Gaspé Current. These waters have lower buffering capacities to decreasing pH and $\Omega_{A}$. Furthermore, the accumulation of metabolic carbon dioxide produced by rapid oxic degradation of organic matter in the sandy and permeable sediments of the Magdalen Shelf or in the Gaspé current could decrease $\Omega_{A}$ in the Southern Gulf (Starr and Chassé, *personal communication*).

We thus divided the and their analysis between the Southern and Northern St. Lawrence. The St. Lawrence was divided along the 350m isobath on the southern slope of the Laurentian Channel, from the Cabot Strait to the tip of the Gaspé Peninsula. The $\Omega_A$ distribution in the Southern St. Lawrence was modeled
using the exponential kriging model.

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. Whereas
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentrations in the water column available through
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2019] and interpolated oxygen % saturation using cokriging with depth as a covariable, as done in @dutil2011. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia data layer
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

Based on the wide range of $\Omega_A$ levels at which organisms are negatively
affected (see above), we built an index of acidification stress ($A_s$)
ranging between 0 and 1 that slowly increases as $\Omega_A$ decreases to $\approx$ $2$, increases more rapidly close to $\Omega_A$ $= 1$ and reaches its peak at
$\Omega_A$ $\approx$ $0.5$:

$$A_s = \frac{-3}{.99 + e^{-2 * \Omega_A}} + 3$$

Hence, the higher the acidification stress, the higher the index of
acidification becomes.

### Hypoxia

The data used to characterize hypoxia come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] survey in late spring and fall of 2013 to 2017, as well as from DFO’s annual multispecies surveys for the northern Gulf in August and for the Magdalen Shallows in September. We provide a brief summary of data and methods to describe hypoxia in this document. For more details, refer to @blais2019.

Oxygen concentration were measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on a CTD. The probe is calibrated against seawater
samples taken directly from the Niskin bottles at every cast and analyzed by Winkler titration [see @blais2019 for calibration procedure]. The data used is the last depth sampled on the CTD profile, which is typically ~10m above the
bottom. Note that this depth does not necessarily correspond to the depth of the oxygen minimum. The oxygen minimum is typically found along the 27.25 isopycnal, often situated as much as 100 m above bottom in the Lower Estuary. Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation, taking into account the salinity and temperature at the selected
depth.

We interpolated oxygen saturation using cokriging with depth as a covariable,
as done in @dutil2011. Bathymetric data come from @dutil2011 and have a
100 $km^2$ resolution. According to @diaz1995, severe hypoxia occurs
when dissolved oxygen falls below 2 $ml$ $L^{-1}$, to 62.5 $\mu mol$ $L^{-1}$ or ~20% saturation. This is considered as the level
necessary to maintain most animal life. Accordingly, @chabot2008 studied the
effects of hypoxia on the energy budget of cod in the St. Lawrence and found
that behavioural effects began manifesting themselves below 70% oxygen saturation and that
survival becomes jeopardized below ~20% saturation. We used these observations
to create an index of hypoxia using an inverted logistic curve that slowly
increases below 70% oxygen saturation and increases acutely as it reaches its peak intensity between 20 and 30% oxygen saturation:

$$H_s = \frac{-1}{.99 + 200 * e^{-0.15 * O_2}} + 1$$

Hence, the higher the hypoxic stress, the higher the index of hypoxia becomes.

### Bottom-water temperature anomalies

The data used to characterize bottom-water temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods to
characterize the bottom-water temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

Bottom-water temperatures are interpolated in the Gulf using
conductivity-temperature-depth (CTD) sampling performed annually through DFO's
multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. Using this sampling survey, temperatures are horizontally
interpolated at each 1 m depth layer on a 2 km resolution grid. Bottom-water
temperatures are then extracted by using a bathymetry layer from the Canadian
Hydrographic Survey [@dutil2012] and selecting the interpolated temperature
from the layer corresponding to the bottom depth at each grid point.

We used temperature anomalies, *i.e.* deviations from long-term normal
conditions between 1981 and 2010, to measure an annual index of stress associated with extreme
temperatures between 2013 and 2017. Temperature anomalies were calculated using
the difference between grid cell values with 1981-2010 climatological averages.
Anomaly time series were normalized by their standard deviation (SD) to allow
comparisons between areas of the St. Lawrence with different temperature ranges.
For example, temperatures observed in deep channels are less variable than in
shallower regions of the St. Lawrence. Hence, if anomalies were expressed in
degrees Celsius, it would underestimate the relative importance of anomalies
in deep channels when compared to shallower regions. Grid cells whose monthly
value exceeded ±0.5 standard deviation (SD) from the long-term average were
considered as anomalous [@galbraith2018]. Outliers in the data were defined as
those that fell beyond the interquartile range * 3, identified as extreme
outliers by @tukey1977. Outlier values were capped to correspond to the 5th and
95th percentile values. Anomalies were divided into positive and negative anomalies and the absolute value of anomalies were used as an indicator of the intensity of bottom-water temperature anomalies. The mean anomaly intensity between 2013 and 2017 for each grid cell was used to generate the final index of bottom-water temperature anomalies.

### Surface-water temperature anomalies

The data used to characterize surface-water temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods to
characterize the surface-water temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

The surface-water layer is characterized using sea surface-water temperature (SST) monthly
composites from Advanced Very High Resolution Radiometer (AVHRR) satellite
images obtained from the National Oceanic and Atmospheric Administration (NOAA)
and European Organization for the Exploitation of Meteorological Satellites
(EUMETSAT). Images used are from DFO’s Maurice Lamontagne Institute at a 1 km
resolution from 1985-2013 and from DFO’s Bedford Institute of Oceanography (BIO)
Operational Remote Sensing group at a 1.5 km resolution since 2014. Monthly
anomalies were constructed as the difference between monthly averages and the
1985-2010 climatological mean for each month.

Surface-water temperature anomalies were characterized following the same method used
for bottom-water temperature anomalies. Only the months of May to November were
included to avoid biases associated with the presence of ice cover. Monthly
anomalies from May to November values were summed to obtain an indicator of
annual surface-water temperature anomaly intensity in each grid cell. The mean
anomaly intensity between 2013 and 2017 for each grid cell was used to generate
the final index of surface-water temperature anomalies.

### Sea-level rise

The data used to characterize sea-level rise risk come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Sea-level rise was characterized by @nicholls2010 using NASA's satellite
altimetry data (Topex/Poseidon, Jason-1&2, GFO, ERS-1&2, and Envisat
missions) and available at
http://www.aviso.altimetry.fr/en/data/products/ocean-indicatorsproducts/mean-sea-level/products-images.html

The rate of sea-level rise ($mm/year$) was measured between 1992 and 2012 and
transformed as a net change value ($mm$) by multiplying by the number of years
considered. Only positive values were selected under the assumption that
only positive sea-level rise is likely to cause environmental stress.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

## Coastal

### Aquaculture

As aquaculture sites are managed at the provincial level in the St. Lawrence, data on aquaculture sites come from various sources
[@mapaq2016; @dfo2016a; @aaf2016; @fa2016; @ffa2016].
Invertebrates aquaculture is especially important in the southern and western
Gulf. Fish and algae aquaculture, on the other hand, remains marginal.
Considering this, we only included invertebrates aquaculture for the
aquaculture driver layer. However, if fish or algae farming were to become more
important, these drivers should be incorporated in future analyses as individual
layers, as impacts vary between types of aquaculture.

Aquaculture activities are highly localized and potential effects do not or
rarely extend beyond the location of the farms. We therefore only considered
the location of sites to characterize the distribution of this driver.
We were unable to characterize site production in terms of biomass farmed, which
could provide an indication of the intensity of aquaculture activities. As such,
we considered aquaculture as binary presence-absence data.

### Coastal development

Terrestrial stable lights at night mostly represent light from human settlements and industrial sites with electricity. We thus used lights at night as a proxy of coastal infrastructure development.

The data come from the Nighttime Lights Time Series. Nighttime light products
are compiled by the Earth Observation Group at the National Oceanic and
Atmospheric Administration's (NOAA) National Centers for Environmental
Information (NCEI). They use globally available nighttime data obtained from
the Visible Infrared Imaging Radiometer Suite (VIIRS) Day/Night Band (DNB) of
the Defense Meteorological Satellite Program (DMSP) to characterize global
average radiance ($nanoWatts$ $cm^{-2}$ $sr^{-1}$) composite images at a
15 arc-second (~200 m) resolution.

We used the annual Version 1 Nighttime VIIRS DNB composites between 2015 and
2016 [@earthobservationgroup2019] to characterize coastal development in coastal areas of the
St. Lawrence. As the effects of coastal development are likely acute in its
direct vicinity, we extracted average radiance values using a 2 km buffer
around grid cells within 2 km of the coast. We used a weighted area average to
extract the radiance values.

### Direct human impact

As in @halpern2008a and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statistics-canada2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated. They combine
to cover all of Canada and are highly variable in shapes and sizes.
\textcolor{blue}{For example, their sizes range from 0.0003 to over 44000 $km^2$ in our study area.}
The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

As the effects of direct human impacts are likely acute mostly in coastal areas
we calculated total population in grid cells within 2 km of the coast. Total
population was measured in a 10 km buffer around each coastal cell.
The total population in each buffer was the sum of intersecting dissemination
areas divided by the intersection area between buffers and dissemination areas:

$$DHI_j = \sum_{k=1}^{n_j} P_k * \frac{A_{j,k}}{A{tot, k}}$$

where $j$ is a buffered grid cell, $k$ is a dissemination area intersecting $j$,
$P$ is the population in $k$, $A$ is the area of the $k$ overlapping with $j$
and $A_{tot}$ is the total area of $k$. This approach was favoured to reduce the
effects of very large dissemination areas overlapping with buffers on a very
small percentage of their total area.

### Inorganic pollution

The data used to characterize inorganic pollution come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Inorganic pollution was modeled using impervious surface area (*i.e.*
artificial surfaces such as paved roads) under the assumption that most of this
pollution source comes from urban runoff. Inorganic pollution originating from
point-sources or in areas lacking paved roads is therefore not captured by
this layer. The data obtained was aggregated at the watershed scale. Spread into coastal and marine environments was modeled using a diffusive plume model
from each watershed pourpoints (*e.g.* river mouths).

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

### Nutrient pollution

The data used to characterize nutrient pollution come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual fertilizer use in tonnes ($t$) was used as a proxy of nutrient pollution.
The data used came from the Food and Agriculture Organization of the United
Nations (FAO). Gaps in data were modeled using a linear regression between
fertilizer and pesticides or agricultural gross domestic product (GDP).
Dasymetric maps were then used to distribute fertilizer data over the landscape
using 2009 data from the Moderate Resolution Imaging Spectroradiometer (MODIS)
at ~500 m resolution and aggregated to watersheds. Diffusive plume models
from each watershed pourpoint (*e.g.* river mouths) were then used to model the
distribution and intensity of nutrient pollution in coastal and marine
environments.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

### Organic pollution

The data used to characterize organic pollution come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual pesticide use in tonnes ($t$) was used as a proxy of organic pollution.
The data used came from the FAO and gaps in data were modeled using a linear
regression between pesticides and fertilizers or agricultural GDP. The same
methodology as that used to characterized nutrient pollution was then applied
to organic pollution.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

### Toxic algae

The data we use to describe the risk of toxic algae come from an expert based
map delineating the areas where coastal areas are at risk from five different
toxins [@bates2019]. The map presents coastal areas at risk from 5
different toxins: 1) paralytic shellfish poisoning (PSP) toxins from the regular
presence of the dinoflagellate *Alexandrium catenella* (previously
*Alexandrium tamarense*) at high concentrations, 2) amnesic shellfish
poisoning (ASP) toxins from domoic acid 3) diarrhetic shellfish poisoning (DSP)
toxins, 4) spirolides and 5) pectenotoxins, two toxins produced by
dinoflagellates occurring in the St. Lawrence.

The information provided on this expert map on the 5 toxins [@bates2019].
was georeferenced and transformed as vectorized objects.
We calculated a toxic algae risk ($T$) index for each cell ($x$) in the 1 $km^2$
study grid. For each toxin ($t$), a value of 1 was attributed to all grid cells
overlapping with areas identified at risk on the expert map and a value of 0.5
for grid cells overlapping with areas where ASP and DSP toxins were observed
without exceeding legal thresholds. The value for all 5 toxins was them summed
for all grid cells:

$$TA_{i,x} = \sum_{i = 1}^{5} i_x$$

## Fisheries

The impacts of fisheries activities in the St. Lawrence are evaluated using
DFO's fisheries logbooks program [@dfo2016]. While logbooks are not mandatory
for all fisheries in the St. Lawrence, they still provide a very thorough
overview of the spatial distribution and intensity of fishing activities. The data we used span 6 years from 2010 to 2015 and detail
218323 fishing events (36387 $\pm$ 3147 fishing events per year). There were 31
targeted species and a total of 53 species caught in the dataset.

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008a and
@halpern2015 and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

Gear types can be further classified into fixed or mobile engines based
on their mobility. We used these two mobility classes to generate a buffer of
impact around each fishing activity coordinates to consider potential spatial
uncertainty associated with locations, the fact that mobile engines can be
tracted over several kilometers during fishing activities, and because we do not
have the beginning and end points of mobile fishing events. Buffer sizes for
fixed and mobile engine was of 200 and 2000 meters, respectively.

Table S2. Classification of gear types in the fisheries dataset based on their
environmental impact and mobility

Gear type (EN)    | Classification    | Mobility
--------------    | --------------    | --------
Trap              | DNH               | Fixed
Trawl             | DD                | Mobile
Dredge            | DD                | Mobile
Driftnet          | PHB               | Fixed
Hand lines        | PLB               | Fixed
Longline          | PHB               | Fixed
Scuba diving      | DNL               | Fixed
Purse seine       | PLB               | Fixed
Seine             | DNH               | Fixed
Beach seine       | DNH               | Fixed
Trap              | DNH               | Fixed
Jig fishing       | PLB               | Fixed

To characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multiplied the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

$$FI_j = \sum_{k=1}^{n_j} B_{tot, k} * \frac{A_{j,k}}{A_{tot,k}}$$

where $j$ is a grid cell, $k$ is a fishing event, $B_{tot}$ is the total biomass
of a fishing event $k$, $A$ is the area of a fishing event *k* overlapping a
cell $j$ and $A_{tot}$ is the total area of the fishing event $k$. This formula
gives an intensity measurement in biomass units, which is kg in our case. Since
we measure the intensity within a 1 $km^2$ grid cell, the intensity evaluation
is in $kg * km^{-2}$. This metric distributes the biomass captured within each
grid cell as a function of overlapping fishing area and provides an overview of
how impacted each grid cell is in terms of extracted biomass.

## Marine traffic

### Shipping

The data used to characterize shipping come from the global cumulative impacts
assessment on habitats [@halpern2008a; @halpern2015a] and available on the NCEAS
online data repository [@halpern2015]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

Two data sources were used to characterize shipping. The first set of data is
gathered as part of the World Meteorological Organization Voluntary Observing
Ships' (VOS) scheme. Ships participating in the program gather
meteorological data along with observation location as part of an open-ocean
climate dataset. The data spans 20 years and annually covers 10-20% of ships
worldwide. Data used spanned 2003 to 2011.

The second set of data comes from the Automatic Identification System (AIS),
an initiative launched in 2002 that sought to improve marine safety by providing
mariners with real-time vessel traffic. Through the
International Maritime Organization SOLAS agreement, all vessels of over 300
gross tonnage on international voyages and those carrying passengers are now
required to be equipped with AIS transceivers. These transceivers use Global
Positioning System technology to locate vessels every 10 minutes. The data used
were from November 2010 to December 2011.
Data used come mostly from vessels that move globally (*i.e.* cargo, tanker and
passenger), as they are required to carry AIS transceivers, but also include
data from fishing, high-speed, pleasure and support classes. Shipping intensity
was evaluated as the number of fishing tracks at a 0.1 decimal degrees resolution.
For more details on data and methods used, consult [@walbridge2013].

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

### Invasive species

The data used to characterize invasive species risk come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
supplementary materials provided in @halpern2008a and @halpern2015.

Cargo volume was used as a proxy of invasion risk under the assumption that
risk of invasion is proportional to tonnes of goods transferred through ports.
Cargo throughput in metric tonnes for the year 2011 was accessed through a
variety of sources [see supplementary material in @halpern2015 for more details]
and cross-matched with entries in the World Port Index database (WPI; available
from the National Geospatial-Intelligence Agency). A gap-filling procedure
using linear regression and sets of predictors related to port volume and
available in the WPI dataset was then applied to the WPI dataset to predict
missing cargo volume entries. Finally, volume data was distributed in marine
environments adjacent to ports using a diffusive plume model with an exponential
decay function that set the maximum spread distance to approximately 1000 km.
The plume model was then clipped to areas less than 60 m deep, as
invasive species are more likely to invade shallow areas.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

### Marine pollution

The data used to characterize marine pollution risk come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015].
Marine pollution was considered to be mainly driver by the shipping industry.
As such, the driver layer was constructed by combining the shipping (*i.e.*
shipping lanes) and invasive species (*i.e.* cargo volume) layers.
invasive. For more details, refer to supplementary materials provided in
@halpern2008a and @halpern2015.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

# Driver intensity and distribution

We evaluated the frequency distribution of each drivers to verify whether data
should be transformed (Figure S\ref{hist}). In light of this, we log-transformed
the following driver layers

- Coastal development
- Direct human impact
- All fisheries data
- Hypoxia
- Inorganic pollution
- Invasive species
- Nutrient pollution
- Organic pollution
- Bottom-water temperature anomalies
- Shipping

To allow for relative intensity comparison, all driver layers were subsequently
normalized between 0 and 1 using the 99th quantile to further control for
extreme values (Figure S\ref{drivers}).

\begin{figure}[H]
\centering
\includegraphics{./FigureS1.png}
\caption{Frequency distribution of the untransformed data for all driver layers.}
\label{hist}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics{./FigureS2.png}
\caption{Distribution and intensity of transformed and normalized drivers in the Estuary and Gulf of St. Lawrence available on *eDrivers*.}
\label{drivers}
\end{figure}

# Cumulative exposure

\begin{figure}[H]
\centering
\includegraphics{./FigureS3.png}
\caption{Frequency distribution of cumulative exposure (\textit{i.e.} sum of normalized driver intensity in each grid cell) and percent contribution of each driver to the frequency distribution of cumulative exposure in the Estuary and Gulf of St. Lawrence.}
\label{marimekko}
\end{figure}

# ~~Threat complexes~~ \textcolor{blue}{Cumulative exposure profiles}

## Clustering

\begin{figure}[H]
\centering
\includegraphics{./FigureS4.png}
\caption{Validation procedure for the \textit{k-medoids} and \textit{k-means} clustering algorithms based on the number of cluster that maximizes average silhouette width (upper panels; Kaufman and Rousseeuw, 1990) and minimizes the total within-cluster sum of squares (WSS; lower panels).}
\label{valid}
\end{figure}

## Inter-cluster dissimilarity

\begin{figure}[H]
\centering
\includegraphics{./FigureS5.png}
\caption{Evaluation of inter-cluster dissimilarity using a similarity percentage analysis (SIMPER) with Manhattan distance (Clarke, 1993). The figure diagonal presents the distribution of the 6 clusters identified using the \textit{k-medoids} clustering algorithm. The lower triangle shows all combinations of inter-cluster dissimilarity with circular barplots showing the percent contribution to total dissimilarity of each driver and with the total inter-cluster dissimilarity in the center of the barplots. The upper triangle shows the average relative intensity of each driver for all driver combinations, with barplots to the left and the right representing the row and columns clusters, respectively.}
\label{inter}
\end{figure}

## Intra-cluster similarity

\begin{figure}[H]
\centering
\includegraphics{./FigureS6.png}
\caption{Evaluation of intra-cluster similarity using the Manhattan distance transformed to a similarity index. The distribution of the 6 clusters is presented along with circular barplots showing the percent contribution to total similarity of each driver and with the total intra-cluster similarity in the center of the barplots.}
\label{intra}
\end{figure}

# References